Three Problems for Sophistication

### 2015-01-21 Facticity/Facticity/paper.tex
...

\title{Two Problems for Sophistication}

@@

\section{Introduction}
Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring. There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity. For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed. In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory. We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.

The Kolmogorov complexity of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but, as shown by a central result, known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the result only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.

@@

But fundamental problems arise as soon as hard cut-offs are introduced. In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue. This is dangerous: if such ad-hoc ``fixes'' result in a theory that is hard to prove either wrong or right, then this can create an artificial dead end for a valuable area of research. When hard cut-offs are avoided on the other hand, all such measures are suddenly no longer necessary. What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.

>>>>

\title{Three Problems for Sophistication}

@@

\section{Introduction}
Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring. There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity. For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed. In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory. We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.

The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.

@@

But fundamental problems arise as soon as hard cut-offs are introduced. In our opinion, a lot of measures taken in the literature to address problems with sophistication, such as restricting the model class or introducing model penalties, complicate the method and make problems harder to analyse, without actually addressing the fundamental issue. This is dangerous: if such ad-hoc fixes result in a theory that is hard to prove either wrong or right, then this can create an artificial dead end for a valuable area of research. When hard cut-offs are avoided on the other hand, all such measures are suddenly no longer necessary. What remains is a simple and elegant theory that can be used to sift through all possible models, disproving most while retaining a select number of interesting candidates for our further consideration.
